Reinforcement Learning Produces Dominant Strategies for the Iterated Prisoner's Dilemma

### 2017-06-04 main.tex
Mostly citations

Key points:
* We trained several different strategies and they all are tournament
winners
* works for standard, noisy, and moran
* Complex strategies can fare well
* the best strategies are cooperative yet punitive
* copious references to W. and D. Ashlock

Todo:
* better title

@@

* Update on Axelrod Library
Description of strategy space and tournament capabilities
Includes many prior tournament winners (OmegaTFT, ZDGTFT2)
Some original strategies but mostly from the literature

@@

@@

Two strategies in the library, Winner12 and Winner21, from \cite{Mathieu2015},
are based on lookup tables for $n_1 = 0$, $m_1 = 1$, and $m_2=2$. The strategy
Winner12 emerged in less than 10 generations of training in our framework using
a score maximizing objective.

@@

This strategy type can be used to train arbitrary memory-$n$ strategies and a
memory one strategy was trained and is called PSO Gambler Mem 1, with
probabilities $\ldots$. Though it performs well in deterministic tournaments
it is not as good as the longer memory strategies, and is bested by a similar
strategy that also uses the first round of play.

% TODO: quantify how stochastic
% TODO: include memory one four vector

@@

% TODO: Cite earlier ANN papers including Ashlock
% TODO: list features

@@

We used strategies based on finite state machines to create a number of
strategies. These strategies are deterministic and are efficient computational.
At each state the machine transitions to a new state and plays a specified move
depending on the opponent's last action. These strategies can encode a variety
of other strategies, including classic strategies like TitForTat, encode
handshakes, and gruding strategies that always defect after an opponent
defection.

% TODO: Cite Ashlock FSM papers
% TODO: One or more diagrams

@@

Finally, we note that as the library grows, the top performing strategies
sometimes shuffle, and are not retrained regularly. The precise parameters
that are optimal will depend on the pool of opponents. Moreover we have not
extensively trained strategies to determine the minimum parameters that are
sufficient -- neural networks with fewer nodes and features and finite state
machines with fewer states may suffice.
% TODO: Cite Ashlock paper on parameter capacity for trained strategies

>>>>

@@

* Update on Axelrod Library

The Axelrod library \cite{knight2016open} now contains over 200 strategies,
many from the scientific
literature, including classic strategies like Win Stay Lose Shift
\cite{nowak1993strategy} and
previous tournament winners such as OmegaTFT \cite{slany2007some},
 Adaptive Pavlov \cite{li2007design}, and ZDGTFT2 \cite{stewart2012extortion}.

@@

There are several previous publications that use evolutionary algorithms to
evolve IPD strategies in various circumstances \cite{fogel1993evolving} \cite{ashlock2006training}
\cite{ashlock2006changes} \cite{vassiliades2010multiagent} \cite{ashlock2014shaped}
\cite{ashlock2014evolution} \cite{ashlock2015multiple} \cite{barlow2015varying}
\cite{sudo2015effects}. See also \cite{gaudesi2016exploiting} for a strategy trained to win against
a collection of well-known IPD opponents. Our results are unique in that we are
able to train against a large collection of well-known strategies available in
the scientific literature.

@@

Two strategies in the library, Winner12 and Winner21, from \cite{Mathieu2015},
are based on lookup tables for $n_1 = 0$, $m_1 = 1$, and $m_2=2$. The strategy
Winner12 emerged in less than 10 generations of training in our framework using
a score maximizing objective. Strategies nearly identical to Winner21 arise
from training with a Moran process objective.

@@

This strategy type can be used to train arbitrary memory-$n$ strategies and a
memory one strategy was trained and is called PSO Gambler Mem 1, with
probabilities $(\text{Pr(C|CC)}, \text{Pr(C|CD)}, \text{Pr(C|DC)}, \text{Pr(C|DD)}) = (1, 0.5217, 0, 0.121)$. Though it performs well in deterministic tournaments
it is not as good as the longer memory strategies, and is bested by a similar
strategy that also uses the first round of play.

These strategies are trained with a particle swarm algorithm rather than an
evolutionary algorithm (though the former would suffice). Particle swarm
algorithms have been used to trained IPD strategies previously
\cite{franken2005particle}.

% TODO: quantify how stochastic

@@

% TODO: list features

@@

We used strategies based on finite state machines to create a number of
strategies. These strategies are deterministic and are efficient computational.
At each state the machine transitions to a new state and plays a specified move
depending on the opponent's last action. These strategies can encode a variety
of other strategies, including classic strategies like TitForTat, encode
handshakes, and grudging strategies that always defect after an opponent
defection.

% TODO: One or more diagrams

@@

Finally, we note that as the library grows, the top performing strategies
sometimes shuffle, and are not retrained regularly. The precise parameters
that are optimal will depend on the pool of opponents. Moreover we have not
extensively trained strategies to determine the minimum parameters that are
sufficient -- neural networks with fewer nodes and features and finite state
machines with fewer states may suffice. See \cite{ashlock2013impact} for
discussion of resource availability for IPD strategies.
